---
title: "Lab-03"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Linear Regression

### Part: 1

First we need to input the dataset. Here we create two vectors with both sets of observations
```{r}
distance <-c(3.4, 1.8, 4.6, 2.3, 3.1, 5.5, .7, 3.0, 2.6, 4.3, 2.1, 1.1, 6.1, 4.8, 3.8)

damage <- c(26.1, 17.8, 31.3, 23.1, 27.5, 36.0, 14.1, 22.3, 19.6, 31.3, 24.0, 17.3, 43.2, 36.4, 26.1)

```

Then we put both into a dataframe to show the relationship
```{r}
fd <- data.frame(distance, damage)
```

Now we can build the model - and display details (gradient & intercept)
```{r}
fit <- lm(damage ~ distance)

fit
```

### Part: 2

We can now plot the two vectors and use the regression line from the model to show their relationship
```{r}
plot(distance, damage)
abline(fit)
```

### Part: 3
```{r}

```

### Part: 4

To check how well the model has fitted to the data, we can use the summary command
```{r}
summary(fit)
```
This shows that the RSE is 2.316. This indicates that prediction for fire damage would on average be 2.316 off when using the least squares line.

The R-squared is 0.9176. This indictates that 91.76% of the sample varaiation can be explained by distance.

### Part: 5

From the model summary we can see that the t-value is 12.525, and the p-value is 1.248e-08. As the t-value is large and the p-value is small, we can conclude that there is a relationship between fire damage and distance to the nearest fire station. 

### Part: 6

We can use the predict function to predict value for new test data
```{r}
newData <- c(0.5, 1.5, 2.5, 3.5)

predict(fit,data.frame(distance=newData))
```

### Part: 7 

We can use the confit function to check confidence level parameters
```{r}
confint(fit, level=0.95)
```

### Part: 8

We can use the predict function, with the additional parametr interval to get predictions with confidence values
```{r}
predict(fit, data.frame(distance=c(0.5, 3, 5)), interval = "confidence")
```


### Part: 9

We can use the same process as in part 8 for prediction intervals 
```{r}
predict(fit,data.frame(distance=(c(0.5, 3, 5))), interval="prediction")
```

### Part: 10

From the results in parts 8 & 9, we can see that prediction intervals are always larger than confidence intervals

### Part: 11

To plot bothconfidence intervals & prediction intervals on the same graph, we can use the following commands 
```{r}
plot(distance, damage, xlab="distance", ylab = "damage", main = "Least Squares Line with Confidence intervals and prediction intervals")
abline(fit)
plotLine <- data.frame(distance=seq(0,6.5,length=65)) # dummy plotline to show intervals 
predict_c = predict(fit,plotLine,interval="confidence")
predict_p = predict(fit,plotLine,interval="prediction")
lines(plotLine$distance,predict_c[,"lwr"],col="firebrick", type="b",pch="+")
lines(plotLine$distance,predict_c[,"upr"],col="firebrick", type="b",pch="+")
lines(plotLine$distance,predict_p[,"upr"],col="midnightblue", type="b",pch="*")
lines(plotLine$distance,predict_p[,"lwr"],col="midnightblue",type="b",pch="*")

```
