---
title: "Lab-04.Rmd"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Logistic Regression 

## Part 1:

First we load the data and get the basic info for the dataset. We can use the summary() command
```{r}
mydata <-read.csv("https://stats.idre.ucla.edu/stat/data/binary.csv")

summary(mydata)

```

Alternatively we can use the view() command 

```{r}
#view(mydata)
```


## Part 2:

To check the number of observations, we can use the nrow command 
```{r}
nrow(mydata)
```

## Part 3:


We can use sapply to check the standard deviation of the variables in the dataset. We exclude rank as it is a catgorical variable.
```{r}
sapply(mydata [,-4], sd)
```

Likewise we can use sapply for the mean values
```{r}
sapply(mydata [,-4], mean)
```

## Part 4: 

```{r}
mydata$rank <- factor(mydata$rank)

mydata$rank
```


## Part 5:

To fit a logistic model to the data, we can use the glm (generalised linear model) command 
```{r}
glm.admit.fit = glm(admit ~ gre + gpa + rank, data = mydata, family = "binomial")

```

Now we can use the summary() command to get details about the model
```{r}
summary(glm.admit.fit)
```

## Part 6:

This process is standard One-hot encoding


## Part 7:

All the variable z-values are large, and all the variable p values are small (<0.05) hence all the variable are statistically significant  


## Part 8:

To obtain the raw probability predictions for the training data, we can use the following command
```{r}
admit.prob = predict(glm.admit.fit, type = "response")
```

We can check the values by using the head() command, which will display the first n rows
```{r}
head(admit.prob)
```


## Part 9:

To convert the probabilities to class, first we create a new vector of lenght 400 (to represent each of the 400 observarions)
```{r}
admit.pred = rep(1,400)
```

We then use the probability data, can use a simple conditional to assign it a class. In this class if a probability if 0.5 or above for a class, it will be assigned to the class.
```{r}
admit.pred[admit.prob<0.5]=0 
```

We can again used the head command to show the data
```{r}
head(admit.pred)
```


## Part 10:

We can use the table command to create a confusion matrix 
```{r}
table(admit.pred, mydata$admit)
```


## Part 11:

First we construct a new dataframe with the mean gre and gpa values for each ranks
```{r}
newdata1 <- with(mydata, data.frame(gre = mean(gre), gpa = mean(gpa), rank = factor(1:4)))

newdata1
```

We can then fit the linear model to the mean value to compute raw probabilities 
```{r}
newdata1$admit1.prob = predict(glm.admit.fit, newdata = newdata1, type = "response")

newdata1
```

Finally we normalise the probabilities to predict a class, using the same process we used in part 9
```{r}
newdata1$admit1.pred = rep(1,4)

newdata1$admit1.pred[newdata1$admit1.prob<0.5]=0

newdata1
```