---
title: "Lab-01"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- First we learnt how to create and set values to variables in R. -->

[//]:(lalal lal alk)


## 1. Vectors 

### Part a

Here we create a vector of 11 consecutive numbers starting from -10. There are several ways this can be accomplished 

```{r}
seq(-10,0)

seq(-10,0, length.out = 11)

c(-10, -9, -8, -7, -6, -5, -4, -3, -2, -1, 0)
```

### Part b

Here we create a vector of 11 numbers, starting from -0.1 with intervals of 0.5. We can do this one of several ways

```{r}
seq(-0.1, by=0.5, length.out = 11)

seq(-0.1, 4.9, length.out = 11)

c(-0.1, 0.4, 0.9, 1.4, 1.9, 2.4, 2.9, 3.1, 3.9, 4.4, 4.9)
```

### Part c

Here we calulate the sum and products of the vectors created in parts a&b

```{r}
u <- seq(-10,0)

v <- seq(-0.1, by=0.5, length.out = 11)
```
Sum
```{r}
u + v
```
Product
```{r}
u * v
```

### Part d
Here we add 1 to all the elements in vector u

```{r}
u + 1
```

And we subtract 20% from all the elements in vector v

```{r}
v * 0.8
```


### Part e
Here we concatanate vectors u & v

```{r}
w <- c(u, v)

w
```


And we report the length of the new vector
```{r}
length(w)
```


### Part f

Here we return a new vectors, containing the 14th, 15th & 16th elements in vector w

```{r}
w[c(14, 15, 16)]
```

Similarly this vectors is formed of the 2nd, 5th, 9th & 21st elements of w

```{r}
w[c(2, 5, 9, 21)]
```

As w only contains 22 element this comment to return the 23rd element will return nothing
```{r}
w[23]
```


### Part g

This command reassign the value the 3rd element in w to 100
```{r}
w[3] <- 100

w
```

This command replaces the values in the 7th, 15th & 22nd positions with new values (200, 300, 400)
```{r}
replace(w, c(7, 15, 22), c(200, 300, 400) )
```

### Part h

This command removes vector u from the enviroment
```{r}
rm(u)
```

### Part i

This command clears the entire enviroment
```{r}
rm(list = ls())
```

## 2. Matrices

### Part a

Here we create a new matrix of dimensions 4x5, starting with the value 1 & incrementing each new value by 2.
The matrix will be populated by row
```{r}
B <- matrix(seq(1, by=2, length.out = 20), 4, 5, byrow = TRUE)

B
```

This command selects element that are in both the 2nd & 4th row & the 2nd and 3rd column

```{r}
B[c(2,4), c(2,3)]
```

### Part b

This function creates a new vector of length n, and populates it with the Fibonacci sequence, start with the value 1
```{r}
Fib <- function(n) {
  if (n == 1) return(1)
  x <- c(1, 1)
  while (length(x) < n) {
    index <- length(x)
    new <- x[index] + x[index - 1]
    x <- c(x,new)
  }
  return(x)
}
```

Now we uses the function as a parameter to create a new matrix of dimentions 3x3

```{r}
C <- matrix(Fib(9), 3,3)

C
```

Here we select the elements not present in the both the 2nd and 3rd row & the 2nd & 3rd column. This returns a new vector with the element in position 1,1 of C

```{r}
C[ -c(2, 3), -c(2, 3)]

```

This command will return a new vector with the element in 1,1 & 1,2

```{r}
C[ -c(3), -c(2 ,3)]
```

This returns a new vector with all elements from the first row of C

```{r}
C[ , -c(2, 3)]
```
