---
title: "Lab-02"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
[//]:(lalal lal alk)

## Basic Stats


### 1. Data Sampling


Here we create a vector of 20 random values 
```{r}
x <- c(10, 2, 15, 6, 4, 9, 12, 11, 3, 0, 12, 10, 9, 7, 11, 10, 8, 5, 10, 6)

x
```

Here we get the number of elements in the vector
```{r}
length(x)
```


This command sums the numbers in the vector 

```{r}
sum(x)
```


This command gives the mean of the numbers in the vector
```{r}
mean(x)
```


This one gives the median
```{r}
median(x)
```

This command computes the mode of the vector
```{r}
names(sort(-table(x)))[1]
```


Here we extract the Quantile Ranges from the vector (Min, Q1, M, Q3, Max)
```{r}
quantile(x)
```

Here we compute the variance

```{r}
var(x)
```


This command computes the standard deviation
```{r}
sd(x)
```


### 2. Distrbutions

Here we return a new vector y, with a normal distribution, with 20 observations, the same mean as vector x, and a standard deviation of 1
```{r}
y <- rnorm(20, mean(x), 1)

y
```

This command computes the co-varaince of x and y 
```{r}
cov(x, y)
```

This command computes the correlation of x and y
```{r}
cor(x, y)
```

Because we have not set a seed value, the value of y will change everytime y is call. Hence the values of the co-varaince & correlation between x & y will also change everytime. We therefore set a seed value, re-assign value to y, with the same parameters
```{r}
set.seed(63)

y <- rnorm(20, mean(x), 1)

y
```


### 3. Scatter Plots
We now create 2 normal distributions, each with 20 observations, and a standard deviation of 1. Vector z has a mean of 8
```{r}
set.seed(100)

z <- rnorm(20, 8, 1)

z
```

Vector u has a mean of 10
```{r}
u <- rnorm(20, 10, 1)

u
```

This command creates a scatter plots of (x,z) & (x,u) & a legend
```{r}
plot(x,z, xlim=range(x, x), ylim=range(z, u), col="red", main= "Plots of 2 normal distributions of x")
points(x,u, pch='+', col='blue')

legend("topleft",pch=c(1,4), col=c("red", "blue"), c("z", "u"), box.col="black")
```




### 4. Box Plots
Here we are creating box plots of x and y, against the same axis
```{r}
boxplot(x,y, names =(c("x", "y")), col=(c("lightblue","orange")),
        main="Plots of 2 normal distributions", xlab="Distribution Values", ylab="Distrubution Name",
        horizontal=TRUE) 
```


