---
title: "Lab-07"
output:
  html_document: default
  pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Random Forests 

First will will need to load the dplyr library, so we can read in csv files. 
```{r}
library(dplyr)

titanic3 = read.csv("..\\..\\Week08\\titanic3.cSV", header = TRUE)
```

Once the data has been load, We can remove features from the dataset that will not be require for our tree models.
```{r}
titanic3=select(titanic3, -name, -ticket, -boat, -body, -home.dest, -cabin) %>% mutate(embarked = factor(embarked),sex = factor(sex), pclass=factor(embarked))

summary(titanic3)
```



## Part 1: Feature Transformation

We will first need to clean and transform the data so it is optimised for our model. We will initially add a new variable survived01, which is a categorical equivalent of survived. 
```{r}
titanic3$survived01 = as.factor(titanic3$survived)

append(titanic3$survived01, titanic3)
```

## Part 2: Build a forest

To use the random forest model, we will need to load the relevant library.
```{r}
# install package
library(randomForest)
```

Random forest can use only complete records. We will removed any observation that does not have values for all the predictor and target variables. 
```{r}
# remove rows with missing data
titanic.clean <- na.omit(titanic3)
```

## Part 3: Split the data

We can now split the data into training and test sets. 
```{r}
set.seed(1)

# split the data in half

# indices 
train_index <- sample(nrow(titanic.clean), nrow(titanic.clean)/2)

# assign data
train <- titanic.clean[train_index, ]
test <- titanic.clean[-train_index, ]
```

We can also assign a couple of new variables, which will be the expected result from the test set. These can be used to construct the confusion matrix. 
```{r}
# target variable from text set 
actual <- titanic.clean[-train_index, "survived"]

actual01 <- titanic.clean[-train_index, "survived01"]
```

## Part 4: Build random forest model with numeric target variable

We will now build a regression random forest model with survived as the target. We will use all other variables, except survided01, as predictors.
```{r}
rf.titanic <- randomForest(survived ~ . -survived01, data = train, mtry = 7,importance=TRUE)
```

We can now used the model to predict probabilities. 
```{r}
prediction <- predict(rf.titanic, newdata = test)
```

We can compare the predicted and actual results to compute the test error.
```{r}
# confusion matrix for test error
mean((prediction - actual) ^ 2)
```

Additionally we can get some stats on the model. 
```{r}
# stats
rf.titanic
```

## Part 5: Part 5: Random forest model with categorical variable as target

Now We will now build a classification random forest. We will use survided01 as the target, and all other variables, with the execption of survived, being used as predictors.

```{r}
rf.titanic01 <- randomForest(survived01 ~ . -survived, data = train, mtry = 7,importance=TRUE)
```

We can now used the model to predict probabilities, with the test dataset. 
```{r}
prediction01 <- predict(rf.titanic01, newdata = test)
```

We can also get model statistics, directly from the model as it is a classification model.
```{r}
# stats
rf.titanic01

# number of tree
rf.titanic01$ntree

# error
mean(rf.titanic01$err.rate)
```

## Part 6: Variable Importance Plot

We can analysis and plot the importance of each predictor variable in a random forest. For the regression model:
```{r}
# Regression Model

importance(rf.titanic)

varImpPlot(rf.titanic)
```

For the classification model:
```{r}
# Classification Model

importance(rf.titanic01)

varImpPlot(rf.titanic01)
```


## Part 7: Comparison of Error Rates 

We can compare the error rate from different modelling approaches. For this comparison, we will use a forest comprising of 200 tree.The first will use a majority vote from all the tree in the forest to decided a classification.  
```{r}
# The black line (madjority vote)

# vector for test error
error.mv = rep(0,200)

for (i in 1:200){
 
  # new classification model
  rf.titanicC <- randomForest(survived01 ~ . -survived, data = train, mtry = 7,importance=TRUE)
  
  # get predictions
  predictionC <- predict(rf.titanicC, newdata = test)
  
  error.mv[i] <- mean(rf.titanicC$err.rate)
}
```

The second approach uses averages probabilities from each tree and predicts the class with the highest value.
```{r}
# blue line 

# vector for test error
error.ave = rep(0,200)

for (i in 1:200){
  
  # new classification model
  rf.titanicR <- randomForest(survived ~ . -survived01, data = train, mtry = 7,importance=TRUE)
  
  # get predictions
  predictionR <- predict(rf.titanicR, newdata = test)
  
  error.ave[i] <- mean((predictionR - actual) ^ 2)
}
```

We can now plot all both of the results against the error from a single tree.
```{r}
# black line
plot(error.mv, xlab="Number of Data sets", ylab="Test Error Rate ", type="l"
     , ylim= c(0.2, 0.3))

#blue line
lines(error.ave, col="blue")

# add red line
abline(h=error.mv[1],lty=2,col="red")
```



## Part 8: Experiments with number of predictors at tree splits

We can now experiment with different values for mtry, to try and reduce the error. 
```{r}
error.preds <- rep(0,7)
```

We will build 7 different forests, each with a different number of possible splits (1-7) at each branch. 
```{r}
for (i in 1:length(error.preds)){
  
  rf.titanicM <- randomForest(survived ~ . -survived01, data = train, mtry = i,importance=TRUE)
  
  predictionM <- predict(rf.titanicM, newdata = test)
  
  error.preds[i] <- mean((predictionM - actual) ^ 2)
  
}
```

We can plot the results.
```{r}
plot(error.preds, type="b", xlab="mtry", ylab="Test MSE")
```


## Part 9: Errors vs ntrees 

We will now experiment by altering the number of tree in each forest, and compare these with different values for mtry. We can visualise the results with a plot.
```{r}
plot(0,ylab="Test Error", xlab="Number of Trees", ylim = c(0.1,0.25), xlim= c(0,500), main="Random Forest")
```

```{r}
errors.trees <- matrix(nrow=7, ncol=500 )
```

```{r}
for (i in 1:7){

  for (j in 1:500) {

    rf.titanicLoop <- randomForest(survived ~ . -survived01, data = train, mtry = i,ntree = j ,importance=TRUE)    
    
    predictionLoop <- predict(rf.titanicLoop, newdata = test)
    
    errors.trees[i, j] <- mean((predictionLoop - actual) ^ 2)
    
    lines(errors.trees[i,],col=i)
    
  }
}
```

```{r}
legend(title="mtry","topright",c("1","2","3","4","5","6","7"),lty=rep(1,7),col=1:7)
```
